---
title: "Regression Models with R Applications - Rahul's Notes"
author: "Rahul"
date: "14/10/2020"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, intro.css]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = FALSE)
```

# Notes from Session
* Email of Dr. Olga
  * Olga.Korosteleva@csulb.edu
* Book has ~22 basic regressions and 60 variations
* Greek letters used 
  * alpha
  * beta
  * epsilon
  * lambda
  * pi
  * sigma
---
# General Linear Regression
* x is fixed as it is observed
* y hence carries the distribution of error and mean determined by the regression coefficients and x
* hat notation is used for estimates
  * beta hat estimated with MLE (max likelihood estimator)
  * for complete information also estimate the variance
---
# Interpretation of fitted coefficients
* How much does a variable influence the response
* x1 is continuous
  * beta 1 is the change in estimated mean of y for a unit change for x1 with all other vraiables remaining same
* x1 is binary variable
  * beta 1 is the change in estimated mean of y as x1 shifts from 0 to 1 while all other variables remain same. 
  * (0 -1) can be viewed as comparison for Male and Female and estimated y is change as x1 changes from Male to Female
---
# Predicted Response
* Estimate output variable

---
* Example Job Satisfaction
* If there is more than three levels in a variable, then dummy variable is created
* So beta hat for that variable is a comparison for change from one to another
  * Male to Female
  * Bachelor to Master
  * Base to other category - done using relevel

* Prediction
  * for bachelor, both education terms disaapear
  * if female, the gender term disappears
---
```{r}
df<- read.csv(file=here::here("exercises","NormalExerciseData.csv"), header=TRUE, sep=",", stringsAsFactors = TRUE)

df <- df %>% mutate(nmeds = as_factor(nmeds))
#age","gender","ethnicity","BMI","nmeds","AQI","HR
#plotting histogram with fitted normal density
library(rcompanion)
plotNormalHistogram(df$HR)

#fitting general linear model
summary(fitted.model<- glm(HR ~ gender + age + ethnicity + BMI + AQI + nmeds, 
data=df, family=gaussian(link=identity)))

#estimating sigma
print(sigma.hat<- sigma(fitted.model))

#using fitted model for prediction
print(predict(fitted.model, data.frame(gender="M", age=50, ethnicity="Hispanic",BMI = 20,AQI = "moderate",nmeds = "0")))



```

# Solve example

---
# Gamma

* right skewed. long right tail
* need all positive variable
  * can add constant to make all positive
* Estimated mean is alpha times beta
* link function - how does the 
* log link function
* need to estimate beta and alpha coefficients

* Interpretation
  * Percent change for unit change in e raised to response minus 1
  * For 0-1 variable no need to subtract 1
  
* Examples of Gamma
  * anything to do with money
  * house prices


---
```{r}

df<- read.csv(file=here::here("exercises","GammaExerciseData.csv"), header=TRUE, sep=",", stringsAsFactors = TRUE)

#rescaling variables



#plotting histogram with fitted normal density
library(rcompanion)
plotNormalHistogram(df$score)

colnames(df) %>% glue::glue_collapse(sep='","')

# "desgn","wrkyrs","priorQI","score"
#fitting gamma regression 
summary(fitted.model<- glm(score ~ desgn + wrkyrs 
+ priorQI , data=df,
family=Gamma(link=log)))


#using fitted model for prediction
print(predict(fitted.model, type="response", data.frame(desgn="nurse",
wrkyrs=7,priorQI="yes")))
```

---
# Binary Logistics Regression

* pi is estimated probability of y=1
* Logit link function
* Logit is natural log of odds

* Interpretation
  * x continuous
    * % change in odds for unit change
  * x 0-1 variable
    * percent ration of odds
---
```{r}

```

---
# Poisson Regression

* Log link function
* output 0,1,2,3,...
* theoretical upto infinite
* lambda is mean expected output
* Interpretation
  * x continuous
    * exponential of (percent change minus 1)
  * x 0-1
    * exponential of (percent change)
    
---

```{r}
hospitalstay.data<- read.csv(file="examples/PoissonExampleData.csv", header=TRUE, sep=",", stringsAsFactors = TRUE)

#fitting Poisson model
summary(fitted.model<- glm(days ~ gender + age + illness, 
data=hospitalstay.data, family=poisson(link=log)))

#using fitted model for prediction
print(predict(fitted.model, data.frame(gender="M", age=55, illness="no"), 
type="response"))
```





  
